Supplementary Information for: Crosslinguistic Word Orders Enable an Efficient Tradeoff between Memory and Surprisal

### 2020-03-29 writeup/si.tex
code

\usepackage{showlabels}
\usepackage{fullpage}
\usepackage{pslatex}
%\usepackage{latexsym}
\usepackage[english]{babel}
\usepackage[utf8]{inputenc}
\usepackage{amsmath}
\usepackage{bm}
\usepackage{graphicx}
\usepackage{tikz}
\usepackage{xcolor}
\usepackage{url}
%\usepackage[colorinlistoftodos]{todonotes}
\usepackage{rotating}
\usepackage{natbib}
\usepackage{amssymb}

@@

@@

>>>>

\usepackage{showlabels}
\usepackage{fullpage}
\usepackage{pslatex}
\usepackage{hyperref}
%\usepackage{latexsym}
\usepackage[english]{babel}
\usepackage[utf8]{inputenc}
\usepackage{amsmath}
\usepackage{bm}
\usepackage{graphicx}
\usepackage{tikz}
\usepackage{xcolor}
\usepackage{url}
%\usepackage[colorinlistoftodos]{todonotes}
\usepackage{rotating}
\usepackage{natbib}
\usepackage{amssymb}

@@

\tableofcontents

@@

\subsection{Results: Japanese}

\begin{figure}
	\begin{center}
		\includegraphics[width=0.3\textwidth]{../code/ngram-control/create_models_ngrams/morph/Japanese/tradeoffs/figures/Japanese-suffixes-byPhonemes-it.pdf}
		\includegraphics[width=0.3\textwidth]{../code/ngram-control/create_models_ngrams/morph/Japanese/tradeoffs/figures/Japanese-suffixes-byPhonemes-memsurp.pdf}
		\includegraphics[width=0.3\textwidth]{../code/ngram-control/create_models_ngrams/morph/Japanese/tradeoffs/figures/Japanese-suffixes-byPhonemes-auc-hist.pdf}
\end{center}
	\caption{Japanese verb suffixes, measuring prediction on the level of phonemes, for real (blue), random (green), and approximately optimized (red) orderings. Left: $I_t$ as a function of $t$. Center: Memory-surprisal tradeoff. Right: Areas under the curve for the memory-surprisal tradeoff.}\label{fig:jap-phon}
\end{figure}

\begin{figure}
	\begin{center}
		\includegraphics[width=0.3\textwidth]{../code/ngram-control/create_models_ngrams/morph/Japanese/tradeoffs/figures/Japanese-suffixes-byMorphemes-it.pdf}
		\includegraphics[width=0.3\textwidth]{../code/ngram-control/create_models_ngrams/morph/Japanese/tradeoffs/figures/Japanese-suffixes-byMorphemes-memsurp.pdf}
		\includegraphics[width=0.3\textwidth]{../code/ngram-control/create_models_ngrams/morph/Japanese/tradeoffs/figures/Japanese-suffixes-byMorphemes-auc-hist.pdf}
\end{center}
	\caption{Japanese verb suffixes, measuring prediction on the level of morphemes, for real (blue), random (green), and approximately optimized (red) orderings. Left: $I_t$ as a function of $t$. Center: Memory-surprisal tradeoff. Right: Areas under the curve for the memory-surprisal tradeoff.}\label{fig:jap-morph}
\end{figure}

\begin{figure}
\begin{center}
\begin{tabular}{c||llll}
             &       Pairs & Full \\ \hline\hline
Optimized for Phoneme Prediction   &   0.976 (SD 0.007) & 0.971 (SD 0.011) \\
Optimized for Morpheme Prediction  &   0.873 (SD 0.154) & 0.85 (SD 0.184) \\
Random Baseline     &  0.519 (SD 0.177) & 0.559 (SD 0.202) \\
\end{tabular}
\end{center}
\caption{Accuracy of approximately optimized orderings, and of random baseline orderings, in predicting verb suffix order in Japanese. `Pairs' denotes the rate of pairs of morphemes that are ordered correctly, and `Full' denotes the rate of verb forms where order is predicted entirely correctly. We show means and standard deviations over different runs of the optimization algorithm (`Optimized'), and over different random orderings (`Random').}\label{fig:acc-japanese}
\end{figure}

\subsection{Results: Sesotho}

\begin{figure}
	\begin{center}
\includegraphics[width=0.3\textwidth]{../code/ngram-control/create_models_ngrams/morph/Sesotho/tradeoffs/figures/Sesotho-suffixes-byPhonemes-it.pdf}
\includegraphics[width=0.3\textwidth]{../code/ngram-control/create_models_ngrams/morph/Sesotho/tradeoffs/figures/Sesotho-suffixes-byPhonemes-memsurp.pdf}
\includegraphics[width=0.3\textwidth]{../code/ngram-control/create_models_ngrams/morph/Sesotho/tradeoffs/figures/Sesotho-suffixes-byPhonemes-auc-hist.pdf}
\end{center}
	\caption{Sesotho verb affixes, measuring prediction on the level of phonemes, for real (blue), random (green), and approximately optimized (red) orderings. Left: $I_t$ as a function of $t$. Center: Memory-surprisal tradeoff. Right: Areas under the curve for the memory-surprisal tradeoff.}\label{fig:jap-phon}
\end{figure}

\begin{figure}
	\begin{center}
\includegraphics[width=0.3\textwidth]{../code/ngram-control/create_models_ngrams/morph/Sesotho/tradeoffs/figures/Sesotho-suffixes-byMorphemes-it.pdf}
\includegraphics[width=0.3\textwidth]{../code/ngram-control/create_models_ngrams/morph/Sesotho/tradeoffs/figures/Sesotho-suffixes-byMorphemes-memsurp.pdf}
\includegraphics[width=0.3\textwidth]{../code/ngram-control/create_models_ngrams/morph/Sesotho/tradeoffs/figures/Sesotho-suffixes-byMorphemes-auc-hist.pdf}
\end{center}
	\caption{Sesotho verb affixes, measuring prediction on the level of morphemes, for real (blue), random (green), and approximately optimized (red) orderings. Left: $I_t$ as a function of $t$. Center: Memory-surprisal tradeoff. Right: Areas under the curve for the memory-surprisal tradeoff.}\label{fig:jap-morph}
\end{figure}

%cp ~/memory-surprisal/code/ngram-control/create_models_ngrams/morph/Japanese/tradeoffs/figures/* figures/
%cp ~/memory-surprisal/code/ngram-control/create_models_ngrams/morph/Sesotho/tradeoffs/figures/* figures/

\begin{figure}
\begin{tabular}{cc||ll|ll}
             &              & \multicolumn{2}{c}{Prefixes}    & \multicolumn{2}{|c}{Suffixes} \\
             &              & Pairs & Full & Pairs & Full \\ \hline\hline
             Phonemes   &   Optimized  &  0.987 (SD 0.004) & 0.983 (SD 0.006) & 0.914 (SD 0.131) & 0.906 (SD 0.14) \\
           &   Random  &  0.468 (SD 0.321) & 0.364 (SD 0.353) & 0.459 (SD 0.224) & 0.409 (SD 0.235) \\ \hline
Morphemes  &   Optimized  &  0.997 (SD 0.0) & 0.995 (SD 0.0) & 0.804 (SD 0.0) & 0.721 (SD 0.0) \\
           &   Random  &  0.521 (SD 0.331) & 0.43 (SD 0.349) & 0.468 (SD 0.233) & 0.401 (SD 0.251) \\
\end{tabular}
\caption{Accuracy of approximately optimized orderings, and of random baseline orderings, in predicting verb affix order in Sesotho. `Pairs' denotes the rate of pairs of morphemes that are ordered correctly, and `Full' denotes the rate of verb forms where order is predicted entirely correctly. We show means and standard deviations over different runs of the optimization algorithm (`Optimized'), and over different random orderings (`Random').}\label{fig:acc-sesotho}
\end{figure}
